{\Huge Supersymmetry} \\ \vspace{5mm} Lecture notes for FYS5190/FYS9190 Supersymmetry

### 2025-12-01 superspace.tex
Final commit for the 2025 fall term. This is what we are using on the exam.

\begin{Exercise}
Demonstrate the correctness of the general expression for the left-handed scalar superfield in Eq.~(\ref{eq:leftscalarsuperfield}). {\it Hint:} You may have use for the spinor identities
\[ (\theta\sigma^\mu \bar{\theta})(\theta\sigma^\nu \bar{\theta})=\frac{1}{2}g^{\mu\nu}\theta\theta\bar\theta\bar\theta, \]
\[ \theta\partial_\mu\psi\theta\sigma^\mu \bar{\theta}=\frac{1}{2}\theta\theta\partial_\mu \psi\sigma^\mu\bar{\theta}. \]
\end{Exercise}

\begin{Answer}
We start from
\begin{equation*}
\Phi(y, \theta) = A(y) + \sqrt{2}\theta\psi(y) + \theta\theta F(y),
\end{equation*}
and $y^\mu \equiv x^\mu + i\theta\sigma^\mu \bar{\theta}$. This gives
\begin{eqnarray*}
\Phi(x, \theta, \bar\theta) &=& A(x + i\theta\sigma\bar{\theta}) + \sqrt{2}\theta\psi(x + i\theta\sigma\bar{\theta}) + \theta\theta F(x + i\theta\sigma \bar{\theta}) \\
&=& A(x) + \partial_\mu A(x)i\theta\sigma^\mu \bar{\theta} + \frac{1}{2}\partial_\mu\partial_\nu A(x)(i\theta\sigma^\mu \bar{\theta})(i\theta\sigma^\nu \bar{\theta})+ \sqrt{2}\theta\psi(x) \\
&&+ \sqrt{2}\partial_\mu\theta\psi(x)i\theta\sigma^\mu \bar{\theta} + \theta\theta F(x) \\
&=& A(x) + i\theta\sigma^\mu \bar{\theta}\partial_\mu A(x) - \frac{1}{2}\partial_\mu\partial_\nu A(x)\frac{1}{2}g^{\mu\nu}\theta\theta\bar\theta\bar\theta+ \sqrt{2}\theta\psi(x) \\
&& + i\sqrt{2}\theta\partial_\mu\psi(x)\theta\sigma^\mu \bar{\theta} + \theta\theta F(x) \\
&=& A(x) + i\theta\sigma^\mu \bar{\theta}\partial_\mu A(x) - \frac{1}{4} \Box A(x)\theta\theta\bar\theta\bar\theta+ \sqrt{2}\theta\psi(x)+\frac{i}{\sqrt{2}}\theta\theta\partial_\mu \psi\sigma^\mu\bar{\theta} + \theta\theta F(x),
\end{eqnarray*}
where we have used that $(\theta\sigma^\mu \bar{\theta})(\theta\sigma^\nu \bar{\theta})=\frac{1}{2}g^{\mu\nu}\theta\theta\bar\theta\bar\theta$ and 
$\theta\partial_\mu\psi\theta\sigma^\mu \bar{\theta}=\frac{1}{2}\theta\theta\partial_\mu \psi\sigma^\mu\bar{\theta}$.
\end{Answer}

\begin{Exercise}
Show the vector superfield supergauge transformation properties for the component fields. {\it Hint:} Use the field redefinitions:
\[\lambda(x) \to \lambda(x) -\frac{i}{2}\sigma^\mu\partial_\mu\bar{\phi}(x),\]
\[D(x) \to D(x) - \frac{1}{4}\Box C(x).\]
\end{Exercise}

\begin{Exercise}
Derive the expression for $V_{WZ}^2$ in Eq.~(\ref{eq:V2_WZ}).
\end{Exercise}

>>>>

\begin{Exercise}[label=left_handed_scalar_superield_x]
Demonstrate the correctness of the general expression for the left-handed scalar superfield in Eq.~(\ref{eq:leftscalarsuperfield}). {\it Hint:} You may have use for the spinor identities
\[ (\theta\sigma^\mu \bar{\theta})(\theta\sigma^\nu \bar{\theta})=\frac{1}{2}g^{\mu\nu}\theta\theta\bar\theta\bar\theta, \]
\[ \theta\partial_\mu\psi\theta\sigma^\mu \bar{\theta}=\frac{1}{2}\theta\theta\partial_\mu \psi\sigma^\mu\bar{\theta}. \]
\end{Exercise}

\begin{Answer}
We start from
\begin{equation*}
\Phi(y, \theta) = A(y) + \sqrt{2}\theta\psi(y) + \theta\theta F(y),
\end{equation*}
and $y^\mu \equiv x^\mu + i\theta\sigma^\mu \bar{\theta}$. This gives
\begin{eqnarray*}
\Phi(x, \theta, \bar\theta) &=& A(x + i\theta\sigma\bar{\theta}) + \sqrt{2}\theta\psi(x + i\theta\sigma\bar{\theta}) + \theta\theta F(x + i\theta\sigma \bar{\theta}) \\
&=& A(x) + \partial_\mu A(x)i\theta\sigma^\mu \bar{\theta} + \frac{1}{2}\partial_\mu\partial_\nu A(x)(i\theta\sigma^\mu \bar{\theta})(i\theta\sigma^\nu \bar{\theta})+ \sqrt{2}\theta\psi(x) \\
&&+ \sqrt{2}\partial_\mu\theta\psi(x)i\theta\sigma^\mu \bar{\theta} + \theta\theta F(x) \\
&=& A(x) + i\theta\sigma^\mu \bar{\theta}\partial_\mu A(x) - \frac{1}{2}\partial_\mu\partial_\nu A(x)\frac{1}{2}g^{\mu\nu}\theta\theta\bar\theta\bar\theta+ \sqrt{2}\theta\psi(x) \\
&& + i\sqrt{2}\theta\partial_\mu\psi(x)\theta\sigma^\mu \bar{\theta} + \theta\theta F(x) \\
&=& A(x) + i\theta\sigma^\mu \bar{\theta}\partial_\mu A(x) - \frac{1}{4} \Box A(x)\theta\theta\bar\theta\bar\theta+ \sqrt{2}\theta\psi(x)+\frac{i}{\sqrt{2}}\theta\theta\partial_\mu \psi\sigma^\mu\bar{\theta} + \theta\theta F(x),
\end{eqnarray*}
where we have used that $(\theta\sigma^\mu \bar{\theta})(\theta\sigma^\nu \bar{\theta})=\frac{1}{2}g^{\mu\nu}\theta\theta\bar\theta\bar\theta$ and 
$\theta\partial_\mu\psi\theta\sigma^\mu \bar{\theta}=\frac{1}{2}\theta\theta\partial_\mu \psi\sigma^\mu\bar{\theta}$.
\end{Answer}
%%%

%%%
\begin{Exercise}
Show the vector superfield supergauge transformation properties for the component fields. {\it Hint:} Use the field redefinitions:
\[\lambda(x) \to \lambda(x) -\frac{i}{2}\sigma^\mu\partial_\mu\bar{\phi}(x),\]
\[D(x) \to D(x) - \frac{1}{4}\Box C(x).\]
\end{Exercise}

\begin{Answer}
The goal here is to find the component field transformations for a vector superfield $V$ under a supergauge transformation
\begin{equation*}
    V \to V' = V - i [\Lambda - \Lambda^\dagger],
\end{equation*}
where $\Lambda$ is a scalar superfield. The general form for a vector superfield is,
\begin{equation}
    \begin{aligned}
        V(x, \theta, \bar{\theta}) &= C(x) + \theta^{A}\chi_A(x) + \bar{\theta}_{\dot{A}}\bar{\chi}^{\dot{A}}(x) + \theta \theta M(x) + \bar{\theta}\bar{\theta}M^*(x) \\
        & \quad + (\theta \sigma^\mu \bar{\theta})V_{\mu}(x) + \theta \theta \bar{\theta}_{\dot{A}}\bar{\lambda}^{\dot{A}}(x) + \bar{\theta}\bar{\theta}\theta^A\lambda_{A}(x) + \theta\theta \bar{\theta}\bar{\theta}D(x),
    \end{aligned}
    \label{vektorsuperfelt2}
\end{equation}
and as we saw in Ex.~\ref{left_handed_scalar_superield_x} we can write a scalar superfield as
\begin{align}
    \label{venstrehendt skalarfelt2} 
    \Phi & =A(x) + i\theta\sigma^\mu\bar{\theta}\partial_\mu A(x) - \frac{1}{4}\theta\theta\bar{\theta}\bar{\theta}\Box A(x) + \sqrt{2}\theta\psi(x) - \frac{i}{\sqrt{2}}\theta\theta (\sigma^\mu\bar\theta)^A\partial_\mu \psi_A (x) + \theta \theta F(x),\\
    \label{høyrehendt skalarfelt2}
    \Phi^\dagger &= A^*(x) - i\theta \sigma^{\mu}\bar{\theta}\partial_\mu A^*(x) - \frac{1}{4}\theta\theta\bar{\theta}\bar{\theta}\Box A^*(x) + \sqrt{2}\bar{\theta}\bar{\psi}(x) + \frac{i}{\sqrt{2}}\bar{\theta} \bar{\theta}(\theta\sigma^{\mu})_{\dot{A}}\partial_\mu\bar{\psi}^{\dot{A}}(x) + \bar{\theta}\bar{\theta}F^*(x).
\end{align}

Using $\Lambda = i \Phi$, and Eqs.~\eqref{vektorsuperfelt2}--\eqref{høyrehendt skalarfelt2}, we get
\begin{align}
    V \to V' &= V - i [\Lambda - \Lambda^\dagger] = V - i\Lambda - (i\Lambda)^\dagger \nonumber\\
   [\Lambda = i\Phi]\quad \quad  &= V + \Phi + \Phi^\dagger \nonumber \\
   [\eqref{vektorsuperfelt2}] \quad \quad &=
   C + \theta^{A}\chi_A + \bar{\theta}_{\dot{A}}\bar{\chi}^{\dot{A}} + \theta \theta M + \bar{\theta}\bar{\theta}M^* + (\theta \sigma^\mu \bar{\theta})V_{\mu} + \theta \theta \bar{\theta}_{\dot{A}}\bar{\lambda}^{\dot{A}} + \bar{\theta}\bar{\theta}\theta^A\lambda_{A} + \theta\theta \bar{\theta}\bar{\theta}D \nonumber \\
    [\eqref{venstrehendt skalarfelt2}] \quad \quad     & \quad \, +A + i(\theta\sigma^\mu\bar{\theta})\partial_\mu A - \frac{1}{4}\theta\theta\bar{\theta}\bar{\theta}\Box A + \sqrt{2}\theta^A\psi_{A} - \frac{i}{\sqrt{2}}\theta\theta (\sigma^\mu\bar\theta)^A\partial_\mu \psi_A + \theta \theta F \nonumber \\
    [\eqref{høyrehendt skalarfelt2}] \quad \quad    & \quad \, +A^* - i(\theta \sigma^{\mu}\bar{\theta})\partial_\mu A^* - \frac{1}{4}\theta\theta\bar{\theta}\bar{\theta}\Box A^* + \sqrt{2}\bar{\theta}_{\dot{A}}\bar{\psi}^{\dot{A}} + \frac{i}{\sqrt{2}}\bar{\theta} \bar{\theta}(\theta\sigma^{\mu})_{\dot{A}}\partial_\mu\bar{\psi}^{\dot{A}} + \bar{\theta}\bar{\theta}F^* \nonumber \\
    &= C + A^* + A + \theta^A( \chi_A + \sqrt{2}\psi_A) + \bar{\theta}_{\dot{A}}(\bar{\chi}^{\dot{A}} + \sqrt{2}\bar{\psi}^{\dot{A}}) + \theta \theta (M + F) + \bar{\theta}\bar{\theta}(M^* + F^*) \nonumber \\
    &\quad \, + (\theta \sigma^\mu \bar{\theta})[V_\mu + i \partial_\mu(A - A^*)] \nonumber\\
    &\quad\,  +\theta \theta \bar{\theta}_{\dot{A}}\left( \bar{\lambda}^{\dot{A}} - \frac{i}{\sqrt{2}}\sigma^{\mu B}_{\:\:\,\dot{A}}\partial_\mu \psi_B\right) + \bar{\theta}\bar{\theta}\theta^{A}\left(\lambda_A + \frac{i}{\sqrt{2}}\sigma^{\mu}_{A\dot{B}}\partial_\mu\bar{\psi}^{\dot{B}}\right) \nonumber \\
    & \quad \, + \theta \theta \bar{\theta}\bar{\theta}\left(D -  \frac{1}{4}\Box[A + A^*]\right) 
    \label{superjustér-vektortrafo}\\
    &= C' + \theta^A \chi_A' + \bar{\theta}_{\dot{A}}\bar{\chi}_{\dot{A}}' + \theta \theta M' + \bar{\theta}\bar{\theta}M'^* + (\theta \sigma^\mu \bar{\theta})V_{\mu}' + \theta \theta \bar{\theta}_{\dot{A}}\bar{\lambda}'^{\dot{A}} + \bar{\theta}\bar{\theta}\theta^A\lambda_{A}' + \theta\theta \bar{\theta}\bar{\theta}D'.\nonumber
\end{align}
In the final two equalities we have the transformed component fields and we can compare to the original expression in \eqref{vektorsuperfelt2} to read off how the component fields  transform under a supergauge transformation. We notice that  $\Box[A + A^*]$ makes an appearance together with the $D$-field, and that  $[A + A^*]$ appears with the $C$-field. By redefining $D \to D - \frac{1}{4}\Box C$, the last term of the supergauge transformation has the form 
\begin{align*}
    \theta \theta \bar{\theta}\bar{\theta}D' = \theta \theta \bar{\theta}\bar{\theta}\left(D -  \frac{1}{4}\Box[A + A^*]\right) \xrightarrow[D \to D - \frac{1}{4}\Box C]{}\, & \theta \theta \bar{\theta}\bar{\theta}\left(D - \frac{1}{4}\Box C - \frac{1}{4}\Box[A + A^*]\right) \\
    =\, & \theta \theta \bar{\theta}\bar{\theta}\left(D -  \frac{1}{4}\Box[C + A + A^*]\right) = \theta \theta \bar{\theta}\bar{\theta}(D' + C').
\end{align*}
We can do the same with $\lambda$. Let $\lambda_A \to \lambda_A + \frac{i}{2}\sigma^\mu_{A\dot{B}}\partial_\mu \bar{\chi}^{\dot{B}}$. This is motivated by $\chi_A' = \chi_A + \sqrt{2}\psi_A$. We do the calculation for $\lambda$, but the same holds for  $\bar{\lambda} = \lambda^\dagger$:
\begin{align*}
    \bar{\theta}\bar{\theta}\theta^{A}\lambda_A' = \bar{\theta}\bar{\theta}\theta^{A}\left(\lambda_A + \frac{i}{\sqrt{2}}\sigma^\mu_{A\dot{B}}\partial_\mu \bar{\psi}^{\dot{B}}\right) \xrightarrow[\lambda_A \to \lambda_A + \tfrac{i}{2}\sigma^\mu_{A\dot{B}}\bar{\chi}^{\dot{B}}]{} 
    & \, \bar{\theta}\bar{\theta}\theta^A\left(\lambda_A + \frac{i}{2}\sigma^\mu_{A\dot{B}}\partial_\mu\left[\bar{\chi}^{\dot{B}} + \sqrt{2} \bar{\psi}^{\dot{B}}\right]\right) \\
    =&\, \bar{\theta}\bar{\theta}\theta^A(\lambda_A' + \frac{i}{2}\sigma^\mu_{A\dot{B}}\partial_\mu \bar{\chi}'^{\dot{B}}).
\end{align*}
With these two redefinitions we see that neither $D$ nor $\lambda$ transform under the supergauge transformation. The other fields can be read directly from Eq.~\eqref{superjustér-vektortrafo}. In all we have
\begin{align*}
    C &\to C' = C + A + A^*, \\
    \chi_A &\to \chi_A' = \chi_A + \sqrt{2}\psi_A, \\
    M & \to M' = M + F, \\
    V_\mu &\to V_\mu' = V_\mu + i \partial_\mu(A - A^*),\\
    \lambda_A &\to \lambda_A' = \lambda_A,   \quad \quad [\text{no transformation}] \\
    D &\to D' = D. \quad \quad \, \,\,[\text{no transformation}]
\end{align*}
\end{Answer}
%%%

%%%
\begin{Exercise}[]
Derive the expression for $V_{WZ}^2$ in Eq.~(\ref{eq:V2_WZ}).
\end{Exercise}

\begin{Answer}
The Wess--Zumino gauge is  defined by the component field relations
\begin{align*}
    A + A^* &= -C, \\
    \psi & = -\frac{1}{\sqrt{2}}\chi, \\
    F &= -M.
\end{align*}
This gauge choice removes many terms from the vector superfield. Inserting into Eq.~\eqref{superjustér-vektortrafo}, we find
\begin{align}
    V \xrightarrow[\text{gauge}]{\text{Wess--Zumino}} V_{\text{WZ}} = (\theta \sigma^\mu \bar{\theta})\left[V_\mu + i\partial_\mu(A - A^*)\right] + \bar{\theta}\bar{\theta}\theta^{A}\lambda_A + \theta \theta \bar{\theta}_{\dot{A}}\bar{\lambda}^{\dot{A}} + \theta \theta \bar{\theta}\bar{\theta}D.
\end{align}
Here, all component fields have at least two factors of $\theta$. This is neat, because then $V_{\text{WZ}}^n = 0$ for all $n > 2$. We then calculate $V^2_{\text{WZ}}$. We notice that all terms, except the first, has three $\theta$ or $\bar{\theta}$ factors, and will not survive the squaring. Thus,
\begin{align*}
    V_{\text{WZ}}^2 &\, \,\,\,= (\theta\sigma^\mu\bar{\theta})[V_\mu + i\partial_\mu (A - A^*)](\theta\sigma^\nu \bar{\theta})[V_\nu + i\partial_\nu (A - A^*)] \\
    &\overset{\eqref{theta-sigma-theta-bar-kvadrat}}{=} \frac{1}{2}\theta \theta \bar{\theta}\bar{\theta}[V_\mu V^\mu + i\partial_\mu(A - A^*)V^\mu + V_\mu i \partial_\mu(A - A^*)-\partial_\mu(A - A^*)\partial^\mu(A - A^*)].
\end{align*}
This gives the following compact expression for the exponential of $V_{\text{WZ}}$,
\begin{align}
    e^{V_{\text{WZ}}} &= 1 + V_{\text{WZ}} + \frac{1}{2}V_{\text{WZ}}^2 \nonumber \\
    &= 1 + (\theta \sigma^\mu \bar{\theta})\left[V_\mu + i\partial_\mu(A - A^*)\right] + \bar{\theta}\bar{\theta}\theta^{A}\lambda_A + \theta \theta \bar{\theta}_{\dot{A}}\bar{\lambda}^{\dot{A}} \nonumber \\
    & \quad \,+ \frac{1}{4}\theta \theta \bar{\theta}\bar{\theta}\left[4D + V_\mu V^{\mu} + 2iV_\mu\partial^{\mu}(A - A^*) - \partial_\mu(A - A^*)\partial^{\mu}(A - A^*) \right].
    \label{Eksponensial, Wess-Zumino}
\end{align}
\end{Answer}
